Privacy-Respecting Type Error Telemetry at Scale

### 2023-12-05 programming23/paper.tex
sec5, label RQs

\begin{description}
  \item[How many sessions use type analysis?]
    A mere \pct{10} of sessions (and of telemetry records overall) opt-in to
    type checking~(\cref{f:dataset-overview}).
    Most of these use \mnonstrict{} mode, and fewer than \pct{1}
    use \mstrict{} mode.

  \item[How often do projects contain modules with different analysis modes?]
    Sessions rarely combine analysis modes.
    Only 512 of \nK{340} sessions ($<\pct{0.15}$) switched to modules with different
    analysis modes.
    This number may be a significant underapproximation because we have data
    only on modules that the creator chose to open, but the widespread use
    of \mnocheck{} mode suggests that it is not off by much.

  \item[How often do sessions turn analysis off?]
    Sessions rarely disable type analysis after opting in.
    Only 233 of the \nK{340} sessions ($<\pct{0.07}$)
    switch to a weaker mode.
    One-fourth of the downgrades switch from \mstrict{}
    to \mnonstrict{}, and therefore keep some type analysis.
    The remainder switch from \mstrict{} to \mnocheck{} (\pct{50})
    or from \mnonstrict{} to \mnocheck{} (\pct{25}).
    %% FILL more insight if we match ups and downs together??

    %% upgrade
    %% 519 total sessions
    %% 176 with up
    %% 263 total up
    %% #hash((nocheck . 0) (nonstrict . 44) (strict . 132))
    %%  nocheck : 0.00\%
    %%  strict : 75.00\%
    %%  nonstrict : 25.00\%
    %%%% #hash(((strict nocheck) . 88)
    %%%%       (nonstrict . 0)
    %%%%       (strict . 0)
    %%%%       ((nonstrict nocheck) . 44)
    %%%%       (nocheck . 0)
    %%%%       ((strict nonstrict) . 44))
    %%%%  (strict nonstrict) : 25.00\%
    %%%%  nocheck : 0.00\%
    %%%%  (nonstrict nocheck) : 25.00\%
    %%%%  strict : 0.00\%
    %%%%  nonstrict : 0.00\%
    %%%%  (strict nocheck) : 50.00\%
    %%
    %% downgrade
    %% 515 total sessions
    %% 233 with down
    %% 320 total down
    %% #hash((nocheck . 0) (nonstrict . 64) (strict . 169))
    %%  nocheck : 0.00\%
    %%  strict : 72.53\%
    %%  nonstrict : 27.47\%
    %%%% #hash(((strict nocheck) . 116)
    %%%%       (nonstrict . 0)
    %%%%       (strict . 0)
    %%%%       ((nonstrict nocheck) . 64)
    %%%%       (nocheck . 0)
    %%%%       ((strict nonstrict) . 53))
    %%%%  (strict nonstrict) : 22.75\%
    %%%%  nocheck : 0.00\%
    %%%%  (nonstrict nocheck) : 27.47\%
    %%%%  strict : 0.00\%
    %%%%  nonstrict : 0.00\%
    %%%%  (strict nocheck) : 49.79\%

  \item[For modules that use type analysis (\mstrict{} and \mnonstrict{}):]\leavevmode
    \begin{description}
      \item[- Which errors arise?]
        Syntax errors are extremely common~(over \pct{50},~\cref{t:type-error-count}), 
        followed by type mismatches~(\pct{20} in \mstrict{}),
        arity errors~(\pct{2}),
        and failure to unpack on optional value~(\pct{2}).
        There are several uncommon errors, such as misusing a table operation,
        and a few that never arise: internal and too-complex errors,
        duplicating or swapping a generic parameter, use of a deprecated API,
        and unsafe dynamic property lookup on a class.

@@

  \item[What impact does type analysis have on the number of \FS{} errors?]
    Opting in to type analysis has little impact on \FS{} errors.
    The percent of \FS{} errors in \mnocheck{} code matches the percent of
    sessions using \mnocheck{} mode~(\cref{f:error-by-mode}); whereas,
    if type analysis helped reduce \FS{} errors, \mnocheck{} would own a
    higher percentage of the errors.
    Furthermore, there is no apparent relation between \FS{} errors and edits,
    regardless of analysis mode~(\cref{f:error-changes}).
    A randomly-chosen edit has a 1/3 chance of increasing, decreasing, or
    maintaining the number of \FS{} errors.

>>>>

\begin{description}
  \item[RQ1.]\leavevmode
    \begin{description}
      \item[- How many sessions use type analysis?]
        A mere \pct{10} of sessions (and of telemetry records overall) opt-in to
        type checking~(\cref{f:dataset-overview}).
        Most of these use \mnonstrict{} mode, and fewer than \pct{1}
        use \mstrict{} mode.

      \item[- How often do projects contain modules with different modes?]
        Sessions rarely combine analysis modes.
        Only 512 of \nK{340} sessions ($<\pct{0.15}$) switched to modules with different
        analysis modes.
        This number may be a significant underapproximation because we have data
        only on modules that the creator chose to open, but the widespread use
        of \mnocheck{} mode suggests that it is not off by much.

      \item[- How often do sessions turn analysis off?]
        Sessions rarely disable type analysis after opting in.
        Only 233 of the \nK{340} sessions ($<\pct{0.07}$)
        switch to a weaker mode.
        One-fourth of the downgrades switch from \mstrict{}
        to \mnonstrict{}, and therefore keep some type analysis.
        The remainder switch from \mstrict{} to \mnocheck{} (\pct{50})
        or from \mnonstrict{} to \mnocheck{} (\pct{25}).
        %% FILL more insight if we match ups and downs together??

        %% upgrade
        %% 519 total sessions
        %% 176 with up
        %% 263 total up
        %% #hash((nocheck . 0) (nonstrict . 44) (strict . 132))
        %%  nocheck : 0.00\%
        %%  strict : 75.00\%
        %%  nonstrict : 25.00\%
        %%%% #hash(((strict nocheck) . 88)
        %%%%       (nonstrict . 0)
        %%%%       (strict . 0)
        %%%%       ((nonstrict nocheck) . 44)
        %%%%       (nocheck . 0)
        %%%%       ((strict nonstrict) . 44))
        %%%%  (strict nonstrict) : 25.00\%
        %%%%  nocheck : 0.00\%
        %%%%  (nonstrict nocheck) : 25.00\%
        %%%%  strict : 0.00\%
        %%%%  nonstrict : 0.00\%
        %%%%  (strict nocheck) : 50.00\%
        %%
        %% downgrade
        %% 515 total sessions
        %% 233 with down
        %% 320 total down
        %% #hash((nocheck . 0) (nonstrict . 64) (strict . 169))
        %%  nocheck : 0.00\%
        %%  strict : 72.53\%
        %%  nonstrict : 27.47\%
        %%%% #hash(((strict nocheck) . 116)
        %%%%       (nonstrict . 0)
        %%%%       (strict . 0)
        %%%%       ((nonstrict nocheck) . 64)
        %%%%       (nocheck . 0)
        %%%%       ((strict nonstrict) . 53))
        %%%%  (strict nonstrict) : 22.75\%
        %%%%  nocheck : 0.00\%
        %%%%  (nonstrict nocheck) : 27.47\%
        %%%%  strict : 0.00\%
        %%%%  nonstrict : 0.00\%
        %%%%  (strict nocheck) : 49.79\%
  \end{description}

  \item[RQ2. For modules that use type analysis (\mstrict{} and \mnonstrict{}):]\leavevmode
    \begin{description}
      \item[- Which errors arise?]
        Syntax errors are extremely common~(over \pct{50},~\cref{t:type-error-count}), 
        followed by type mismatches~(\pct{20} in \mstrict{}),
        arity errors~(\pct{2}),
        and failure to unpack on optional value~(\pct{2}).
        There are several uncommon errors, such as misusing a table operation,
        and a few that never arise: internal and too-complex errors,
        duplicating or swapping a generic parameter, use of a deprecated API,
        and unsafe dynamic property lookup on a class.

@@

  \item[RQ3. What impact does type analysis have on the number of \FS{} errors?]
    Opting in to type analysis has little impact on \FS{} errors.
    The percent of \FS{} errors in \mnocheck{} code matches the percent of
    sessions using \mnocheck{} mode~(\cref{f:error-by-mode}); whereas,
    if type analysis helped reduce \FS{} errors, \mnocheck{} would own a
    higher percentage of the errors.
    Furthermore, there is no apparent relation between \FS{} errors and edits,
    regardless of analysis mode~(\cref{f:error-changes}).
    A randomly-chosen edit has a 1/3 chance of increasing, decreasing, or
    maintaining the number of \FS{} errors.
